Oral contraceptives

### 2021-01-12 r_notebooks/exploring_oral_contraceptives.Rmd
Saving smaller .rdsfiles, LPA & k-means analysis

```{r setup}
## Packages
# devtools::install_github("https://github.com/w-hardy/serCymruTools")
library(tidyverse)
library(lubridate)
library(tsibble)
library(rstan) # needed to load prophet
library(prophet)
library(fable)
library(fable.prophet)
library(feasts)
library(furrr)
library(serCymruTools)

@@

## Write to local file
# oc %>%
#   write_rds("../data/oral_contraceptives.rds")

@@

```{r, health-board-modelling}
# system.time(oc_hb_cv_models <-
#   regionalCV(data = oc_monthly_quantity_hb_pre_cv19))
# 
# oc_hb_cv_models %>%
#   write_rds("../data/oc_hb_cv_models.rds")

@@

```{r}
# plan(multisession)
# system.time(oc_hb_fcsts <- regionalJointFcsts(oc_monthly_quantity_hb_pre_cv19))
# plan(sequential)
# 
# oc_hb_fcsts %>%
#   write_rds("../data/oc_hb_fcsts.rds")

@@

```{r, gp-cluster-modelling}
# system.time(oc_gp_cluster_cv_models <-
#   regionalCV(data = oc_monthly_quantity_cluster_pre_cv19))
# 
# oc_gp_cluster_cv_models %>%
#   write_rds("../data/oc_gp_cluster_cv_models.rds")

@@

```{r}
# plan(multisession)
# system.time(oc_gp_cluster_fcsts <- regionalJointFcsts(oc_monthly_quantity_cluster_pre_cv19))
# plan(sequential)
# 
# oc_gp_cluster_fcsts %>%
#   write_rds("../data/oc_gp_cluster_fcsts.rds")

@@

# system.time(oc_gp_cv_models <-
#   oc_monthly_quantity_gp_pre_cv19 %>% 
#     filter(!regional_unit %in% incomplete_gp) %>% 
#               regionalCV())
# 
# oc_gp_cv_models %>%
#   write_rds("../data/oc_gp_cv_models.rds")

@@

```{r}
# plan(multisession)
# system.time(oc_gp_fcsts <-
#               oc_monthly_quantity_gp_pre_cv19 %>%
#               filter(!regional_unit %in% incomplete_gp) %>%
#               regionalJointFcsts())
# plan(sequential)
# oc_gp_fcsts %>%
#   write_rds("../data/oc_gp_fcsts.rds")

oc_gp_fcsts <- 
  read_rds("../data/oc_gp_fcsts.rds")
```

@@

```{r}
library(tidyLPA)
best_model_gp_cluster %>% 
  select(.model, regional_unit) %>% 
  inner_join(mutate(oc_gp_cluster_fcsts,
                    datename = yearmonth(datename)), 
             by = c(".model", "regional_unit")) %>%
  left_join(mutate(select(oc_monthly_quantity_cluster, 
                          c(regional_unit, datename, n)), 
                   datename = yearmonth(datename)),
            by = c("datename", "regional_unit")) %>% 
  mutate(discrepancy = n - .mean,
         z_discrepancy = discrepancy/sd)  %>%
    select(discrepancy, regional_unit, datename) %>%
  pivot_wider(names_from = datename, values_from = discrepancy) %>% 
  ungroup() %>% 
  select(-regional_unit) %>% 
  # estimate_profiles(n_profiles = 4, models = 1) %>% plot_profiles()
    estimate_profiles(1:5,
                      variances = c("equal", "varying"),
                      covariances = c("zero", "varying")
                      ) %>% 
    compare_solutions(statistics = c("AIC", "BIC"))
```

@@

best_model_gp %>% 
  select(.model, regional_unit) %>% 
  inner_join(mutate(oc_gp_fcsts,
                    datename = yearmonth(datename)), 
             by = c(".model", "regional_unit")) %>%
  left_join(mutate(select(oc_monthly_quantity_gp, 
                          c(regional_unit, datename, n)), 
                   datename = yearmonth(datename)),
            by = c("datename", "regional_unit")) %>% 
  mutate(discrepancy = n - .mean,
         z_discrepancy = discrepancy/sd) %>% 
  ggplot(aes(y = discrepancy)) +
  # geom_boxplot(aes(x = as.factor(datename))) +
  geom_line(aes(x = datename, y = discrepancy, colour = as.factor(regional_unit)),
            alpha = .8, lty = 2) +
  geom_hline(aes(yintercept = 0), colour = "red") +
  theme(legend.position = "none")
```

>>>>

```{r setup}
## Packages
devtools::install_github("https://github.com/w-hardy/serCymruTools")
library(tidyverse)
library(lubridate)
library(tsibble)
library(rstan) # needed to load prophet
library(prophet)
library(fable)
library(fable.prophet)
library(feasts)
library(furrr)
library(rebus)
library(MplusAutomation)
library(tidyLPA)
library(serCymruTools)

@@

## Write to local file
# oc %>%
#   saveRDS("../data/oral_contraceptives.rds")

@@

```{r, health-board-modelling}
# plan(multisession)
# system.time(oc_hb_cv_models <-
#   regionalCV(data = oc_monthly_quantity_hb_pre_cv19))
# plan(sequential)
# 
# oc_hb_cv_models %>%
#   saveRDS("../data/oc_hb_cv_models.rds")

@@

```{r}
# plan(multisession)
# system.time(oc_hb_fcsts <- regionalJointFcsts(oc_monthly_quantity_hb_pre_cv19))
# plan(sequential)
# 
# oc_hb_fcsts %>%
#   saveRDS("../data/oc_hb_fcsts.rds")

@@

```{r, gp-cluster-modelling}
# plan(multisession)
# system.time(oc_gp_cluster_cv_models <-
#   regionalCV(data = oc_monthly_quantity_cluster_pre_cv19))
# plan(sequential)
# 
# oc_gp_cluster_cv_models %>%
#   saveRDS("../data/oc_gp_cluster_cv_models.rds")

@@

```{r}
# plan(multisession)
# system.time(oc_gp_cluster_fcsts <- 
#               regionalJointFcsts(oc_monthly_quantity_cluster_pre_cv19))
# plan(sequential)
# 
# oc_gp_cluster_fcsts %>%
#   saveRDS("../data/oc_gp_cluster_fcsts.rds")

@@

# plan(multisession)
# system.time(oc_gp_cv_models <-
#   oc_monthly_quantity_gp_pre_cv19 %>%
#     filter(!regional_unit %in% incomplete_gp) %>%
#               regionalCV())
# plan(sequential)
# 
# oc_gp_cv_models %>%
#   saveRDS("../data/oc_gp_cv_models.rds")

@@

```{r gp-modelling-accuracy}
best_model_gp <- 
  oc_gp_cv_models %>% 
  group_by(regional_unit) %>% 
  arrange(RMSE, MAE) %>% 
  slice(1, .preserve = TRUE)

best_model_gp %>% 
  left_join(select(gp, practiceid, npat, ngp), 
            by = c("regional_unit" = "practiceid")) %>% 
  select(-c(.model, .type)) %>% 
  DataExplorer::plot_correlation()

```{r gp-practice-forecasts}
# plan(multisession)
# system.time(oc_gp_fcsts <-
#               oc_monthly_quantity_gp_pre_cv19 %>%
#               filter(!regional_unit %in% incomplete_gp) %>%
#               regionalJointFcsts())
# plan(sequential)
# 
# oc_gp_fcsts %>%
#   saveRDS("../data/oc_gp_fcsts.rds")

oc_gp_fcsts <- 
  read_rds("../data/oc_gp_fcsts.rds")

oc_gp_fcsts %>% 
  group_by(.model) %>% 
  count

* 2941.64 seconds 2021-01-08

@@

```{r, eval = FALSE}
best_model_gp_cluster %>% 
  select(.model, regional_unit) %>% 
  inner_join(mutate(oc_gp_cluster_fcsts,
                    datename = yearmonth(datename)), 
             by = c(".model", "regional_unit")) %>%
  left_join(mutate(select(oc_monthly_quantity_cluster, 
                          c(regional_unit, datename, n)), 
                   datename = yearmonth(datename)),
            by = c("datename", "regional_unit")) %>% 
  mutate(discrepancy = n - .mean,
         z_discrepancy = discrepancy/sd)  %>%
  select(discrepancy, regional_unit, datename) %>%
  pivot_wider(names_from = datename, values_from = discrepancy) %>% 
  ungroup() %>% 
  select(-regional_unit) %>% 
  estimate_profiles(1:6, models = c(1:3, 6)) %>% 
  compare_solutions(statistics = c("AIC", "BIC"))
```

@@

best_model_gp %>% 
  ggplot(aes(x = MAPE)) + 
  geom_histogram()

best_model_gp %>% 
  group_by(if_else(MAPE < 50, TRUE, FALSE)) %>% 
  count()

median(best_model_gp$MAPE)

acc_filter_value <- 50

best_model_gp_discrep <- 
  best_model_gp %>% 
  filter(MAPE <= acc_filter_value) %>% 
  select(.model, regional_unit) %>% 
  inner_join(mutate(oc_gp_fcsts,
                    datename = yearmonth(datename)), 
             by = c(".model", "regional_unit")) %>%
  left_join(mutate(select(oc_monthly_quantity_gp, 
                          c(regional_unit, datename, n)), 
                   datename = yearmonth(datename)),
            by = c("datename", "regional_unit")) %>% 
  mutate(discrepancy = n - .mean,
         z_discrepancy = discrepancy/sd)

best_model_gp_discrep %>% 
  ggplot(aes(y = discrepancy)) +
  geom_line(aes(x = datename, y = discrepancy, 
                colour = as.factor(regional_unit)),
            alpha = .8, lty = 2) +
  geom_hline(aes(yintercept = 0), colour = "red") +
  theme(legend.position = "none")

```{r}
best_model_gp_discrep %>% 
  filter(discrepancy < -40)

oc_monthly_quantity_gp %>% 
  filter(regional_unit %in% c("W97009", "W98053")) %>% 
  as_tsibble(key = regional_unit, index = datename) %>% 
  autoplot()

gp %>% 
  filter(practiceid %in% c("W97009", "W98053"))
```

* Based on postcodes, these are both University GPs. Or at least, given their procimity to universities, they will likely have large student populations.

#### Latent profile analysis

```{r gp-lpa}
gp_lpa_fit <- 
  best_model_gp_discrep  %>%
  select(discrepancy, regional_unit, datename) %>%
  pivot_wider(names_from = datename, values_from = discrepancy) %>% 
  ungroup() %>% 
  select(-regional_unit) %>% 
  estimate_profiles(1:6, models = c(1:3, 6))

gp_lpa_fit %>% 
  compare_solutions(statistics = c("AIC", "BIC"))

gp_lpa_fit %>% 
  compare_solutions(statistics = c("AIC", "BIC")) %>% 
  .[[1]] %>% 
  ggplot(aes(x = Classes, y = BIC, group = Model, colour = as.factor(Model))) +
  geom_line()

gp_lpa_fit %>%
  plot_profiles(x = ., rawdata = FALSE)

#### mplus

```{r, mplus-data-prep}
best_model_gp_discrep %>% 
  mutate(regional_unit = as.factor(regional_unit)) %>% 
  select(regional_unit, datename, discrepancy) %>% 
  filter(!is.na(discrepancy)) %>% # need to remove NA before pivoting to enable values_fill
  pivot_wider(names_from = datename, values_from = discrepancy, values_fill = -999) %>% 
  left_join(select(gp, c(practiceid, cluster)), by = c("regional_unit" = "practiceid")) %>% 
  mutate(regional_unit = as.factor(regional_unit),
         cluster = as.factor(cluster)) %>% 
  mutate(regional_unit = as.numeric(regional_unit),
         cluster = as.numeric(cluster)) %>% 
  select(cluster, regional_unit, everything()) %>% 
  write_csv(file = "../data/mplus/oc_gp_discrep.csv", col_names = FALSE)

```{r, mplus-anlayses}
source("../r_scripts/functions/mplus_plotting.R")
# mplus_models <- MplusAutomation::readModels("../mplus_analyses/oc_gp")

mplus_model_summaries <- 
  readModels("../mplus_analyses/oc_gp", what = "summaries", quiet = TRUE)

best_mplus_model_summary <- 
  mplus_model_summaries %>% 
  flatten() %>% 
  map(2) %>% 
  bind_rows() %>% 
  filter(Estimator == "MLR" & AnalysisType == "mixture") %>% 
  arrange(BIC) %>% 
  slice(1)

best_mplus_model <- 
  readModels(paste0("../mplus_analyses/oc_gp/", best_mplus_model_summary$Filename))

oc_discrep_5_class_ml <- 
  readModels("../mplus_analyses/oc_gp/oc_discrep_5_class_ml.out")

best_mplus_model$class_counts

# mplus.view.plots("../mplus_analyses/oc_gp/oc_discrep_5_class_ml.gh5")
mplus.plot.estimated_means("../mplus_analyses/oc_gp/oc_discrep_5_class_ml.gh5")
```
#### kmeans

```{r, gp-kmeans}
gp_kmeans_data <- 
  best_model_gp_discrep %>%  
  mutate(regional_unit = as.factor(regional_unit)) %>% 
  select(regional_unit, datename, discrepancy) %>% 
  filter(!is.na(discrepancy)) %>% # need to remove NA before pivoting to enable values_fill
  pivot_wider(names_from = datename, values_from = discrepancy) %>% 
  as_tibble() %>% 
  na.omit()

1:6 %>% 
  map_dfr(~as.data.frame(kmeans(x = gp_kmeans_data[2:6], centers = .x, nstart = 25)$cluster), 
          .id = "n_clusters") %>% 
  mutate(regional_unit = rep(gp_kmeans_data$regional_unit, 6)) %>% 
  pivot_wider(names_from = "n_clusters", id_cols = regional_unit,
              values_from = "kmeans(x = gp_kmeans_data[2:6], centers = .x, nstart = 25)$cluster")

1:6 %>% 
  map_dfr(~as.data.frame(kmeans(x = gp_kmeans_data[2:6],
                                centers = .x, nstart = 25)$centers), 
          .id = "n_clusters") %>% 
  rownames_to_column(var = "cluster") %>% 
  mutate(cluster = 
           str_remove_all(cluster, 
                          pattern = DOT %R% DOT %R% DOT %R% 
                            one_or_more(DIGIT))) %>% 
  pivot_longer(cols = starts_with("2020 "), names_to = "datename") %>% 
  ggplot(aes(x = yearmonth(datename), y = value, colour = as.factor(cluster))) + 
  geom_line() +
  facet_wrap(~n_clusters)

```
#### Comparing methods

```{r}
mplus_assignment <- 
  read_table("../mplus_analyses/oc_gp/oc_discrep_5_class_ml_cprob.txt", col_names = FALSE, na = "*") %>% 
  na.omit() %>% 
  .$X17

kmeans_assignment <- 
  kmeans(x = gp_kmeans_data[2:6], centers = 5, nstart = 25)$cluster

lpa_assignment <- 
  gp_lpa_fit$model_1_class_5$model$classification

df <- bind_cols(mplus_assignment, kmeans_assignment, lpa_assignment)

df %>% 
  ggplot(aes(x = mplus_assignment, y = kmeans_assignment, 
             colour = as.factor(mplus_assignment)))+
  geom_jitter(aes(shape = as.factor(kmeans_assignment)), alpha = .5)

df %>% 
  ggplot(aes(x = mplus_assignment, y = lpa_assignment, 
             colour = as.factor(kmeans_assignment))) +
  geom_jitter(alpha = .5)
```
